Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-02-28 article-models-tt.tex
moved the currying lemma
first proj of x.u is x

@@

\begin{proof}
Let $\jfam{\ctxext{\Gamma}{A}}{P}$ and $\jfam{\ctxext{\Gamma}{A}}{Q}$ be
families. Then we have
\begin{align*}
\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} \tag{by definition}\\
& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} \tag{by \autoref{comp-ww-f}}\\
& \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} \tag{by \autoref{comp-ws-f}}\\
& \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
\end{align*}
\end{proof}

@@

\begin{lem}
When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
More precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
\end{align*}
\end{lem}
Finally, when we substitute by an extended term we get an equal result as when we
substitute two consecutive times. This looks like currying.

>>>>

When we substitute by an extended term we get an equal result as when we
substitute two consecutive times, like the way currying works. Note however,
that we can now \emph{prove} this property using the inference rules we have
already established. It is not a matter of convention, the way it usually is
in type theory by agreeing that $f(x)(y)$ is written as $f(x,y)$.

@@

\begin{proof}
Let $\jfam{\ctxext{\Gamma}{A}}{P}$ and $\jfam{\ctxext{\Gamma}{A}}{Q}$ be
families. Then we have
\begin{align*}
\subst{\cprojfstf{A}{P}}{\ctxwk{\ctxext{A}{P}}{Q}} 
& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{\ctxext{A}{P}}{Q}} \tag{by definition}\\
& \jdeq \subst{\ctxwk{P}{\idfunc[A]}}{\ctxwk{P}{{A}{Q}}} \tag{by \autoref{comp-ww-f}}\\
& \jdeq \ctxwk{P}{\subst{\idfunc[A]}{\ctxwk{A}{Q}}} \tag{by \autoref{comp-ws-f}}\\
& \jdeq \ctxwk{P}{Q} \tag{by \autoref{idfunc-wk-defn}}
\end{align*}
To prove the second judgmental equality, let $\jterm{\Gamma}{A}{x}$ and
$\jterm{\Gamma}{\subst{x}{P}}{u}$. Then we have
\begin{align*}
\cprojfst{A}{P}{\ctxext{x}{u}}
& \jdeq \subst{\ctxext{x}{u}}{\ctxwk{P}{\idfunc[A]}}\\
& \jdeq \subst{u}{{x}{\ctxwk{P}{\idfunc[A]}}} \\
& \jdeq \subst{u}{\ctxwk{\subst{x}{P}}{\subst{x}{\idfunc[A]}}}\\
& \jdeq \subst{x}{\idfunc[A]}\\
& \jdeq x.
\end{align*}
\end{proof}

@@

\begin{lem}
When we substitute an extended term $\ctxext{f}{g}$ of $\ctxext{P}{Q}$ by a term
$x$ of $A$, the term that we get is $\ctxext{\subst{x}{f}}{\subst{x}{g}}$.
More precisely, the following inference rules are valid:
\begin{align*}
& \inference{\jterm{\ctxext{\Gamma}{A}}{P}{f}\quad\jterm{\ctxext{\Gamma}{A}}{\subst{f}{Q}}{g}}{\jtermeq{\Gamma}{\ctxext{\subst{x}{P}}{\subst{x}{Q}}}{\subst{x}{\ctxext{f}{g}}}{\ctxext{\subst{x}{f}}{\subst{x}{g}}}}\\
& \inference{\jterm{\ctxext{{\Gamma}{A}}{P}}{Q}{g}\quad\jterm{\ctxext{{\Gamma}{A}}{P}}{\subst{g}{R}}{t}}{\jtermeq{\ctxext{\Gamma}{\subst{x}{P}}}{\ctxext{\subst{x}{Q}}{\subst{x}{R}}}{\subst{x}{\ctxext{g}{t}}}{\ctxext{\subst{x}{g}}{\subst{x}{t}}}}
\end{align*}
\end{lem}
